---
title: "Patient Health Analysis Using R"
author: "Palak Chavda | Enrollment No.: 2505101200102"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

# Introduction
This Tiny Project analyzes patient health records using R.

**Student:** Palak Chavda  
**Enrollment:** 2505101200102

# Objectives
- Import CSV data
- Clean and inspect the dataset
- Calculate descriptive statistics
- Create graphs
- Analyze correlations
- Interpret results

# Import Data

```{r}
health_data <- read.csv("Palak_Chavda_Patient_Health.csv")
head(health_data)
```

# Data Inspection

```{r}
summary(health_data)
str(health_data)
colSums(is.na(health_data))
sum(duplicated(health_data))
```

# BMI Statistics

```{r}
mean(health_data$BMI)
median(health_data$BMI)
var(health_data$BMI)
sd(health_data$BMI)
```

# Risk Level Distribution

```{r}
risk_count <- table(health_data$Risk_Level)
pie(risk_count, main="Risk Level Distribution")
```

# Average BMI by Gender

```{r}
gender_bmi <- aggregate(BMI ~ Gender, data=health_data, FUN=mean)
barplot(gender_bmi$BMI, names.arg=gender_bmi$Gender,
        main="Average BMI by Gender")
```

# BMI vs Blood Pressure

```{r}
plot(health_data$BMI, health_data$Blood_Pressure,
     pch=19, main="BMI vs Blood Pressure")
abline(lm(Blood_Pressure ~ BMI, data=health_data))
cor(health_data$BMI, health_data$Blood_Pressure)
```

# Findings

- Mean BMI: 28.24
- Median BMI: 28.30
- Standard deviation: 3.89
- BMI and Blood Pressure correlation: 0.952

# Conclusion
This project demonstrates the complete R data-analysis workflow using healthcare data.
